### 2015-07-16 manuscript/ms.tex
add fig  3

\documentclass[12pt,preprint]{aastex} 
%\documentclass[apjl]{emulateapj}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{hyperref}

\newcommand*{\blue}{\textcolor{blue}}

\shortauthors{Kreidberg}

\begin{document}

\title{BATMAN}

\author{Laura Kreidberg\altaffilmark{1,2}}

\email{E-mail: laura.kreidberg@uchicago.edu}

\altaffiltext{1}{Department of Astronomy and Astrophysics, University of Chicago, 5640 S.~Ellis Ave, Chicago, IL 60637, USA}
\altaffiltext{2}{National Science Foundation Graduate Research Fellow}

\begin{abstract}
Batman -- the best super-hero.
\end{abstract}

\keywords{planets and satellites: atmospheres --- planets and satellites:  composition --- planets and satellites: individual: WASP-12b}

\section{Introduction}
%Previous methods --

%Mandel and Agol 2002 - analytic, but uses brute force integration for nonlinear
%Giminez 2006 -
%Pal 2008 - 
%Kipping 2011 -
%Pal 2011 - multiple bodies, possibly lacking radial symmetry
%	basically same as Mandel and Agol for linear/quadratic; more complex LD not implemented; code available (see discussion)
%Abubekerov 
%Kjurkchieva - brute force; radially symmetric; 
%PyTransit

\section{Algorithm}
\label{section:algorithm}
To calculate the fraction $\delta$ of stellar flux blocked by a transiting planet, one must integrate the sky-projected intensity of the star ($I$) over the area obscured by the disk of the planet ($S$):
\begin{equation}
\delta = \iint_S{I \mathrm{d}S}
\label{eqn:surfaceint}
\end{equation}
where $I$ is normalized such that the integrated intensity over the stellar disk is unity. This expression is valid for any general stellar surface brightness map; however, it is slow to evaluate numerically because the differential area elements must be small ($<10^{-6}$) in order to achieve better than one part per million (ppm) accuracy.

On the other hand, if the stellar intensity profile is radially symmetric, the two-dimensional calculation in Equation\,\ref{eqn:surfaceint} can be reduced to one dimension and sped up greatly with the following algorithm: 
\begin{equation}
\delta = \sum_{i=1}^{n} I\left(\frac{x_i+x_{i-1}}{2}\right) \left[A(x_{i}, r_p, d) - A(x_{i-1}, r_p, d)\right]
\end{equation}
where $x$ is the normalized radial coordinate $0 < x < 1$, $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.  The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s$ is the stellar radius.  The intersecting area is given by:

\begin{equation}
A(x, r_p, d) = 
\begin{cases}
x^2\cos^{-1}{u} + r_p^2\cos^{-1}{v} - 0.5\sqrt{w}, & r_p -d < x < r_p + d\\
x^2, & x \le r_p - d\\
r_p^2 & x \ge r_p +d \\
\end{cases}
\end{equation}
where
\begin{eqnarray}
u &=& (d^2+x^2-r_p^2)/(2dx)\\
v &=& (d^2 + r_p^2 -x^2)/(2dr_p) \\
w &=& (-d+x+r_p)(d+x-r_p)(d-x+r_p)(d+x+r_p).
\end{eqnarray}

The advantage of this integration scheme is that the only error introduced is due to approximating the stellar intensity as constant over the differential area element $\Delta A = A(x_i, r_p, d) - A(x_{i-1}, r_p, d)$.  Computation of $\Delta A$ is expensive, but it needs to be calculated relatively few times ($<<10^{6}$ for sub-ppm accuracy). This makes the integration faster than one with a simpler area element (e.g., $\Delta A = \Delta x \Delta y$), that requires a much smaller step size to achieve the same accuracy.

To further optimize the integration, I use a nonuniform step size.  Typical stellar intensity profiles have larger gradients near the limb of the star than at the center \citep[e.g.][]{claret00}, so smaller steps are required at larger $x$ values to achieve the same accuracy.  I adopt the following step-size scaling: 
$$
x_i - x_{i-1}  = f\cos^{-1}\left(x_{i-1}\right)
$$  
where $f$ is a constant scale factor. This prescription is fast to compute and well-behaved at the limits $x=0$ and $x=1$.

See Figure\,\ref{fig:integration} for a schematic illustrating the geometry of the integration.

\begin{figure}
\resizebox{\hsize}{!}{\includegraphics{f1.pdf}}
\caption{Schematic illustration of the integration scheme. The star (large black circle; partially visible) has a radius of 1 and is centered in the plane of the sky at ($x$, $y$) = (0, 0). The planet (smaller black circle) is separated from the center of the stellar disk by a distance $d$ (marked by the solid black line).  The star is partitioned into concentric circles (dotted lines) in order to calculate the integral over the planet disk.  A single integration element $\Delta A$ is shaded in orange. The integration step size illustrated here is larger than for a typical calculation for visual clarity.}
\label{fig:integration}
\end{figure}

\section{The \texttt{batman} package}
I wrote a Python package, \texttt{batman}, to calculate model transit light curves.  \texttt{batman} is an open source project and is being developed on GitHub.  Full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}, but I summarize the main capabilities of the package here.

\subsection{Features}
\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear, quadratic, nonlinear, and custom stellar intensity profiles:
\begin{align}
  I(\mu) &= I_0                                                                         & &\text{(uniform)}             \\
  I(\mu) &= I_0[1 - u_1(1-\mu)]                                                         & &\text{(linear)}              \\
  I(\mu) &= I_0[1 - u_1(1 - \mu) - u_2(1-\mu)^2]                                        & &\text{(quadratic)}           \\
  I(\mu) &= I_0[1 - u_1(1-\mu^{1/2}) - u_2(1- \mu) - u_3(1-\mu^{3/2}) - u_4(1-\mu^2)]   & &\text{(nonlinear)}
\end{align}
where $\mu = \sqrt{1-x^2}$.  A template is available in the source distribution for the creation of custom profiles for any radially symmetric function.  The nonlinear and custom models are calculated with the brute force integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.

For models calculated by integration, the user can specify the maximum allowable truncation error with the \texttt{max\_err} parameter.  When a transit model is initialized, the integration step size is tuned such that the difference between the model and a high-accuracy template (calculated with scale factor $f = 10^{-4}$) is below the specified tolerance.  The lower limit on achievable accuracy with this method is below $10^{-10}$ for typical limb darkening profiles. Figure\,\ref{fig:transit} shows an example transit light curve and its truncation error.

\begin{figure}
\resizebox{\hsize}{!}{\includegraphics{f2.pdf}}
\caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and truncation error for the calculation (bottom panel).  The error tolerance parameter \texttt{max\_err} was set to 1.0 parts per million.  The truncation error increases with distance from the center of the star up to around $\pm0.2$ hours from the time of mid-transit, because the stellar intensity gradient is larger at larger radii. The error decreases again during ingress and egress as the planet eclipses a smaller fraction of the stellar disk.}
\label{fig:transit}
\end{figure}

One of the required inputs for a transit model is the separation of centers $d$ between the star and the planet.  \texttt{batman} computes $d$ automatically from the following set of user-specified parameters: planet semi-major axis, orbital inclination, eccentricity, longitude of periastron, orbital period, and time of periastron/mid-transit time (for eccentric/circular orbits). FIXME add refs to eqns in Seager 10.

The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.  The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation. The number of processors is specified by the user with the parameter \texttt{nthreads}.  \texttt{batman} will raise an exception if the user attempts to parallelize a calculation on a system where OpenMP is not supported.

\subsection{Performance}
\begin{figure}
\resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
\caption{Execution time as a function of truncation error tolerance.}
\label{fig:timing}
\end{figure}

\subsection{Comparison with other codes}

\section{Discussion}

\acknowledgments

\bibliographystyle{apj}
\bibliography{ms.bib}

\end{document}